title: "Portfolio 5: Creating A Random Dataset"
author: "Ryan Wheat"
date: "2023-04-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages}
library(tidyr)
library(dplyr)
```

Background: I need to create a dataframe for my 312 PSYC METHODS students to analyze. 

The Topic: Previous research has found that wearing a backpack depletes mental resources such that objects seem further away when wearing one, than without wearing one. The scenario we seek to test is whether sugar and/or caffeine counteract such an effect. Thus, we have 2 (sugar vs. no sugar) X 2 (caffeine vs. no caffeine) experimental design.

The students are already writing a method section for testing this hypothesis, but they need to write a results section as well -- and in order to do that, they need to have some results to analyze! That's where this portfolio comes into play (and we may do a few other fun things...)

In this dataframe, a "0" indicates that no sugar was given to the participant, and "1" documents that the participant did, in fact, receive sugar as a function of the manipulation. The same pattern holds for caffeine: "0" = no, "1" = yes.

The following chunk creates the dataset by identifying the conditions participants were randomly assigned to, and creating data values randomly sampled from a distribution centered on a mean and standard deviation that I identify. This simulates what an actual experiment is like: these scores are randomly sampled from the "population" distribution (which has been decided by me)!

```{r create-dataframe}

#creating the participants variable, upon which all other values are contingent

participant <- c(1:120)

#create dataframe, and add new variables

backpack_data <- data.frame(participant)

backpack_data <- backpack_data %>%
  mutate(sugar = case_when(participant <= 60 ~ "0",
                           participant > 60 ~ "1"),
         caffeine = case_when(participant > 30 & participant <= 60 ~ "1",
                              participant > 90 & participant <= 120 ~ "1",
                              .default = "0"),
         Distance_Est = c(rnorm(n = 30, mean = 30, sd = 5),
                                             rnorm(n = 30, mean = 40, sd = 8),
                                             rnorm(n = 30, mean = 30, sd = 8),
                                             rnorm(n = 30, mean = 45, sd = 8)))

view(backpack_data)
```

Let's run a 2 X 2 ANOVA on these data to see how the results shake out. Then, we can visualize our results as well.

```{r}
 
#run anova

library(readr)
library(multcompView)

ANOVA <- aov(Distance_Est ~ sugar*caffeine, data = backpack_data)
summary(ANOVA)

```

So we've got a main effect of caffeine, a trend of sugar, and an interaction. Let's plot our results so we can be more specific with this interpretation.

```{r visualize-backpack}



```

